### 2018-04-25 response.tex

\response{
Done. Thanks.
}

\comment{2}{%
Line 11: ``...the objective \textbf{of}...''
}

\response{
Added. Thanks.
}

\comment{3}{%
Lines 25--41: The frequent use of phrases such as ``Among these methods,''; ``Among the latter,''; ``In this category,''; and ``Within this group,'' make the text difficult to follow.
}

% \response{
% \color{red}{Ricardo, please address this comment.}
% }

@@

% \comment{11}{%
% Lines 241--243: It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
% }
%
% \response{
% We added a brief description about them. Thanks.
% }

\comment{12}{%
Lines 257--258: I don't understand this sentence. What is a continuous attribute?
}

\response{
Following Equation (3) we define $A[a]$ as all the possible values of attribute $A$. The use of $[a]$ implies that $A$ is considered a numerically discrete attribute. We therefore meant to say that if $A$ continues attribute. Thanks for the comment. Fixed.
}

\comment{13}{%
Lines 259--264: This paragraph contains a lot of jargon.
}

\response{
Indeed. We expanded the paragraph to include an explanation about the options used with the C5.0 algorithm.
}

@@

\response{
We eliminated ``Similarly''. Thanks.
}

\comment{19}{%
Line 450: ``...remain valid.''
}

\response{
The extra ``s'' was deleted. Thanks.
}

\comment{20}{%
Line 455: ``...as the decisive \textbf{parameters}...'' ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., ``This latter point'', ``One the one hand'', ``on the other hand.''
}

\response{
Indeed. We modified the last paragraph of the Conclusions by removing those sentences.
}

### 2018-04-25 conclusions.tex
update

This latter point is perhaps the most important conclusion of our study. Identifying the total energy and response spectra metrics as the decisive ones is a key contribution to validation of ground motion simulations. On the one hand, this contributes to clarifying a standing question in the area of validation, and on the other hand, it signals to simulators where the focus should be in the search for improvements in their models.

>>>>

\change{Identifying the total energy and response spectra metrics as decisive metrics in the comparisons between synthetics and observations is a key contribution to validation of ground motion simulations. This contributes to clarifying a standing question in the area of validation, and it provides an indication to simulators about where the focus in the search for improvements in their models.}

% This latter point is perhaps the most important conclusion of our study. Identifying the total energy and response spectra metrics as the decisive ones is a key contribution to validation of ground motion simulations. On the one hand, this contributes to clarifying a standing question in the area of validation, and on the other hand, it signals to simulators where the focus should be in the search for improvements in their models.

### 2018-04-25 decision-tree.tex
update

In essence, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome. The process is recursive because new branches and decision nodes are created based on the remaining data at every branch and level, until the tree reaches a certain depth or when other conditions are met. The effectiveness of an attribute $A$ in classifying any subset $S$ from the training data is measured through the information gain $G$ as
% 
\begin{equation}
	\label{eq:gain}
	G(S,A) = E(S) - \sum_{v \, \in \, A[a]} \frac{ \left| S_v \right| }{ \left| S \right| } E \left( S_v \right)
	\, ,
\end{equation} 
% 
where $A[a]$ is the set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
% 
\begin{equation}
	E(S) = \sum_{i=1}^{c} \Big( -p_i \log_2 \left(p_i\right) \Big)
	\, ,
\end{equation}
% 
where $p_i$ is the proportion of $S$ belonging to class $i$, and $c$ are the different values that a given target attribute can take.

@@

In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical. We set the options to do feature selection \myrevision{or winnowing. Therefore, the algorithm choose to use the most important attributes and ignore others. We also activate the evaluation of possible advanced splits of data. Sometimes data is very close to the splitting threshold where assigning a hard threshold may incorrectly change the classification, whereas, that small difference could happen because of numerical errors or error in measurement. Therefore, advanced split can define a soft threshold to consider different probabilities in assigning data to different classes. A decision tree can grow so deep to finally classify all training data correctly. However, the tree will be very complex and will suffer from overfitting. In other words it will have very poor functionality for test data. Therefore, decision tree algorithms (here C5.0) first grow the tree up to the most complex tree then based on user defined factors start to prune the grown tree.The pruning process helps to reduce the complexity of the tree and also increase it's generalizability. Consequently the pruned tree's leaf nodes will have training data which belongs to more than one class which the final class will be the mode of that leaf node. In this study we use two parameters to control the pruning process including confidence or certainty factor (CF) and the smallest number of samples that must be put in at least two of the splits ($S_{\min}$, also known as \textit{minCases}). C5.0 algorithm uses error-based pruning method (see \citet{Quinlan_1993_Book} for more details). The objective is to minimize the observed error rate. Therefore, error estimates for leaf nodes and subtrees are computed assuming that they were used to classify a set of unseen cases with a predicted error rate which is computed based on CF and binomial distribution of number of misclassified events (E) and all training data at that leaf (N). Decreasing CF and increasing $S_{\min}$ cause more heavy pruning process \citep[see][]{Kuhn_2017_Manual}.} \myremove{and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.} In particular, we build a suite of trees for varying values of CF and $S_{\min}$.

>>>>

In \change{general}, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome. The process is recursive because new branches and decision nodes are created based on the remaining data at every branch and level, until the tree reaches a certain depth or when other conditions are met. The effectiveness of an attribute $A$ in classifying any subset $S$ from the training data is measured through the information gain $G$ as
% 
\begin{equation}
	\label{eq:gain}
	G(S,A) = E(S) - \sum_{v \, \in \, A[a]} \frac{ \left| S_v \right| }{ \left| S \right| } E \left( S_v \right)
	\, ,
\end{equation} 
% 
where $A[a]$ is the set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
% 
\begin{equation}
	E(S) = \sum_{i=1}^{c} \Big( -p_i \log_2 \left(p_i\right) \Big)
	\, ,
\end{equation}
% 
where $p_i$ is the proportion of $S$ belonging to class $i$, and $c$ are the different values that a given target attribute can take.

@@

\change{
In the particular case of the C5.0 algorithm, in addition to the general steps just described, the result of the process of building a decision tree also depends on other logical and numerical parameters. We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, \textcolor{red}{(c) discard the use of boosted models,} and (d) use a confidence factor and a threshold in the number of samples that go in th tree leaves. The feature selection option is used by the algorithm to choose the most important attribute over others. The option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes. Discarding the use of boosted models prevents the algorithm from... \textcolor{red}{[pending for discussion with Naeem]}. Last, the options regarding the use of a confidence factor and a minimum number of samples ($S_{\min}$, also known as \textit{minCases}) that can be put in at least two of the branches are set to limit the level of complexity of the tree. In essence, the C5.0 algorithm grows a tree with many branches and then goes through a pruning process to satisfy a minimum number of samples in each leaf and minimize the error of misclassifying the data, which is measure by the confidence factor (CF). Additional details about these options are available in \citet{Quinlan_1993_Book} and \citet{Kuhn_2017_Manual}. For this study, in order to obtain a variety of tree options, we build trees for varying values of CF and $S_{\min}$.
}

% SUGGESTED BY NAEEM:
% In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical. We set the options to do feature selection \myrevision{or winnowing. Therefore, the algorithm choose to use the most important attributes and ignore others. 
% We also activate the evaluation of possible advanced splits of data. Sometimes data is very close to the splitting threshold where assigning a hard threshold may incorrectly change the classification, whereas, that small difference could happen because of numerical errors or error in measurement. Therefore, advanced split can define a soft threshold to consider different probabilities in assigning data to different classes. 
% A decision tree can grow so deep to finally classify all training data correctly. However, the tree will be very complex and will suffer from overfitting. In other words it will have very poor functionality for test data. Therefore, decision tree algorithms (here C5.0) first grow the tree up to the most complex tree then based on user defined factors start to prune the grown tree. The pruning process helps to reduce the complexity of the tree and also increase it's generalizability. Consequently the pruned tree's leaf nodes will have training data which belongs to more than one class which the final class will be the mode of that leaf node. In this study we use two parameters to control the pruning process including confidence or certainty factor (CF) and the smallest number of samples that must be put in at least two of the splits ($S_{\min}$, also known as \textit{minCases}). C5.0 algorithm uses error-based pruning method (see \citet{Quinlan_1993_Book} for more details). The objective is to minimize the observed error rate. Therefore, error estimates for leaf nodes and subtrees are computed assuming that they were used to classify a set of unseen cases with a predicted error rate which is computed based on CF and binomial distribution of number of misclassified events (E) and all training data at that leaf (N). Decreasing CF and increasing $S_{\min}$ cause more heavy pruning process \citep[see][]{Kuhn_2017_Manual}.} In particular, we build a suite of trees for varying values of CF and $S_{\min}$.

% ORIGINAL:
% In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical. We set the options to do feature selection and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models. In particular, we build a suite of trees for varying values of CF and $S_{\min}$.

### 2018-04-25 introduction.tex
update

Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}. \change{Validation entails comparisons of simulations with observations, whereas verification involves comparisons of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}. Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. Among these methods, there are some that are better suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation. Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms. In this category, the comparisons are done both in time and frequency. This type of validation is better suited for high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}). Within this group, methods such as those proposed by \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data in a broad sense.

>>>>

Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}. \change{Validation entails comparisons of simulations with observations, whereas verification involves comparisons of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}. Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. \change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation. Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms. In this category, the comparisons are done both in time and frequency. This type of validation is better suited for high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}). Within this group, methods such as those proposed by \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data in a broad sense.
